---
title: "Jawbone Analysis for Pratical Machine Learning"
author: "Jonathan Garcia Lima"
date: "June 5, 2015"
output:
  html_document:
    pandoc_args: [
      "+RTS", "-K128m", "-RTS"
    ]
---

# Preparing the environment

```{r}
set.seed(5463654)
library(caret)
```

# Loading and cleaning the Data

Now I'll load the data and clean it so that all that's left on my dataset are numeric filled fields.
I'll remove the 7 first fields that are irrelevant to our analysis (those contain data like the user name and the timestamp of the reading).
Afterwards I'll convert all the columns that are factors to numeric and then change all NA columns to zero.

```{r, cache=TRUE}
full <- read.csv("/home/rstudio/pml-training.csv")
oficialTesting <- read.csv("/home/rstudio/pml-testing.csv")
clean <- full[,-c(1,2,3,4,5,6,7)]

n <- ncol(clean)-1
for (i in 1:n) {
  if (is.factor(clean[,i])){
    clean[,i] <- as.numeric(clean[,i])
  }
}

clean <- clean[,colSums(is.na(clean))==0]
```

# Training

I'll start doing a training set with 60% of the data that's already classified so that we can measure if the prediction's good through confusion matrix.

```{r, cache=TRUE}
trainIndex = createDataPartition(clean$classe, p = .6)[[1]]
training = clean[ trainIndex,]
testing = clean[-trainIndex,]

fit <- train(classe ~ ., training, method = "rf")
```

# Confusion Matrix Testing:

```{r, cache=TRUE}
pred <- predict(fit, testing)
conf <- confusionMatrix(pred, testing$classe)
```

Confusion Matrix looks good:

```{r}
conf$table
```

```{r}
conf$byClass
```

Creating a final fit model:

```{r, cache=TRUE}
finalFit <- train(classe ~ ., clean, method = "rf")
```

# Final Prediction:

Applying the same logic to cleanse the oficial testing data:

```{r}
oficialTestingClean <- oficialTesting[, which(names(oficialTesting) %in% names(clean))]

n <- ncol(oficialTestingClean)-1
for (i in 1:n) {
  if (is.factor(oficialTestingClean[,i])){
    oficialTestingClean[,i] <- as.numeric(oficialTestingClean[,i])
  }
}

oficialTestingClean[is.na(oficialTestingClean)] <- 0

finalPred <- predict(finalFit, oficialTestingClean)
```

Generating the files from the file prediction:

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(finalPred)
```